```{r}
mammals_c <- read.csv('mammals_clean.csv')
mammals_g <- read.csv('mammals_genus.csv')
mammals_o <- read.csv('mammals_order.csv')
```

```{r}
# mammals <- Mammal_lifehistories_v2
```

```{r}
library(tidyverse)
```


# Statistical analysis for age of first reproduction

```{r}
afr_result <- lm(genus_size ~ afr_mo, data=mammals_g)
summary(afr_result)
```

# Randomization test
```{r}
set.seed(16)
reshuffled <- mammals_c
reshuffled$afr_mo <- sample(reshuffled$afr_mo, size = nrow(reshuffled), replace = FALSE)
slope_real <- coef(afr_result)["afr_mo"]
simulated_slopes <- list()
nreps = 5000 # 5000 iterations
for(i in 1:nreps){
    
    reshuffled <- mammals_c
    reshuffled$afr_mo <- sample(reshuffled$afr_mo, size = nrow(reshuffled),
                                      replace = FALSE)
    
    # Calculate the slope
    genus_temp <- reshuffled %>%
          group_by(Genus) %>% 
          summarize(genus_size = n_distinct(species), mass_g = mean(mass_g, na.rm = TRUE),
                    gestation_mo = mean(gestation_mo, na.rm = TRUE),
                    newborn_g = mean(newborn_g, na.rm = TRUE),
                    weaning_mo = mean(weaning_mo, na.rm = TRUE),
                    wean_g = mean(wean_g, na.rm = TRUE),
                    afr_mo = mean(afr_mo, na.rm = TRUE),
                    life_mo = mean(life_mo, na.rm = TRUE),
                    litter_size = mean(litter_size, na.rm = TRUE),
                    litters_per_year = mean(litters_per_year, na.rm = TRUE))
    
    reshuffled_res <- lm(genus_size ~ afr_mo, data = genus_temp)
    slope_sim <- coef(reshuffled_res)["afr_mo"]
    
    # Append simulated slope to list
    simulated_slopes[i] <- slope_sim
}    
# Unlist simulated means list into numeric vector
simulated_slopes <- unlist(simulated_slopes)
ggplot() +
    ylab("Count") + xlab("Simulated slope") +
    geom_histogram(aes(x = simulated_slopes), bins = 30, 
                   fill = "grey", alpha = 0.4, colour = "black") +
    geom_vline(xintercept = slope_real, size = 1, 
               linetype = "dashed", colour = "black") + 
    ggtitle("Randomization test for age of first reproduction") + 
    theme_classic()
 ggsave("Randomization test for age of first reproduction.png") 
```
```{r}
abs_simulated_slopes <- abs(simulated_slopes)
exceed_count <- length(abs_simulated_slopes[abs_simulated_slopes >= 
                                               abs(slope_real)])
p_val <- exceed_count / nreps
p_val
```

# Anova

```{r}
anova(lm(genus_size ~ afr_mo*life_mo*gestation_mo*litter_per_year, data=mammals_g))
```


#Linear regression
```{r}
regression_afr <- lm(genus_size ~ afr_mo, data = mammals_g)
ggplot(mammals_g, aes(x = afr_mo, y = genus_size)) +
    geom_point() +
    geom_smooth(method = "lm")
```
```{r}
summary(regression_afr)
```

# Without rodents
```{r}
mammals_c %>%
  filter(order != "Rodentia") %>% 
  group_by(Genus) %>% 
  summarize(genus_size = n_distinct(species), mass_g = mean(mass_g, na.rm = TRUE),
            gestation_mo = mean(gestation_mo, na.rm = TRUE),
            newborn_g = mean(newborn_g, na.rm = TRUE),
            weaning_mo = mean(weaning_mo, na.rm = TRUE),
            wean_g = mean(wean_g, na.rm = TRUE),
            afr_mo = mean(afr_mo, na.rm = TRUE),
            life_mo = mean(life_mo, na.rm = TRUE),
            litter_size = mean(litter_size, na.rm = TRUE),
            litters_per_year = mean(litters_per_year, na.rm = TRUE)) %>% 
  ggplot(aes(x = afr_mo, y = genus_size)) +
    geom_point() +
    geom_smooth(method = "lm") 
  
```


```{r}
genus_no_rod <- mammals_c %>%
  filter(order != "Rodentia") %>% 
  group_by(Genus) %>% 
  summarize(genus_size = n_distinct(species), mass_g = mean(mass_g, na.rm = TRUE),
            gestation_mo = mean(gestation_mo, na.rm = TRUE),
            newborn_g = mean(newborn_g, na.rm = TRUE),
            weaning_mo = mean(weaning_mo, na.rm = TRUE),
            wean_g = mean(wean_g, na.rm = TRUE),
            afr_mo = mean(afr_mo, na.rm = TRUE),
            life_mo = mean(life_mo, na.rm = TRUE),
            litter_size = mean(litter_size, na.rm = TRUE),
            litters_per_year = mean(litters_per_year, na.rm = TRUE))
regression2_afr <- lm(genus_size ~ afr_mo, data = genus_no_rod)
summary(regression2-afr)
# p value is too large
```
# Testing normality by creating the Normal Q-Q plot
```{r}
# png("Normality Test.png")
par(mfrow=c(2,2))
plot(lm(genus_size ~ gestation_mo, data=mammals_g), which=2, xlab = "Gestation time")
plot(lm(genus_size ~ afr_mo, data=mammals_g), which=2, xlab = "Age of first reproduction")
plot(lm(genus_size ~ litters_per_year, data=mammals_g), which=2, xlab = "Litters per year")
plot(lm(genus_size ~ life_mo, data=mammals_g), which=2, xlab = "Life span")
#dev.off()
```

# Testing normality after doing sqrt transformation to x
```{r}
#png("Transformation Test.png")
par(mfrow=c(2,2))
plot(lm(genus_size ~ sqrt(gestation_mo), data=mammals_g), which=2, xlab = "Gestation time")
plot(lm(genus_size ~ sqrt(afr_mo), data=mammals_g), which=2, xlab = "Age of first reproduction")
plot(lm(genus_size ~ sqrt(litters_per_year), data=mammals_g), which=2, xlab = "Litters per year")
plot(lm(genus_size ~ sqrt(life_mo), data=mammals_g), which=2, xlab = "Life span")
#dev.off()
```

# Testing normality after doing higher root transformation to x 
```{r}
#png("Higher Power Normality Test.png")
higer_power_gestation <- mammals_g$gestation_mo ^ (1/4)
higer_power_afr <- mammals_g$afr_mo ^ (1/4)
higer_power_litters<- mammals_g$litters_per_year ^ (1/4)
higer_power_lifespan <- mammals_g$life_mo ^ (1/4)
par(mfrow=c(2,2))
plot(lm(mammals_g$genus_size ~ higer_power_gestation), which=2, xlab = "Gestation time")
plot(lm(mammals_g$genus_size ~ higer_power_afr), which=2, xlab = "Age of first reproduction")
plot(lm(mammals_g$genus_size ~ higer_power_litters), which=2, xlab = "Litters per year")
plot(lm(mammals_g$genus_size ~ higer_power_lifespan), which=2, xlab = "Life span")
#dev.off()
```

# Testing constant variance by creating Scale-Location graph
```{r}
#png("Constant Variance Test.png")
par(mfrow=c(2,2))
plot(lm(genus_size ~ gestation_mo, data=mammals_g), which=3)
plot(lm(genus_size ~ afr_mo, data=mammals_g), which=3)
plot(lm(genus_size ~ litters_per_year, data=mammals_g),which=3)
plot(lm(genus_size ~ life_mo, data=mammals_g), which=3)
#dev.off()
```

```{r}
#png("Higher Power Constant Variance Test.png")
par(mfrow=c(2,2))
plot(lm(mammals_g$genus_size ~ higer_power_gestation), which=3)
plot(lm(mammals_g$genus_size ~ higer_power_afr), which=3)
plot(lm(mammals_g$genus_size ~ higer_power_litters), which=3)
plot(lm(mammals_g$genus_size ~ higer_power_lifespan), which=3)
#dev.off()
```


# Testing normality after dropping 4 orders
```{r}
library(dplyr)
png("Drop 4 orders Randomization Test.png")
no_elephant_whale <- mammals_o %>% 
  filter(order != "Rodentia" & order != "Cetacea" & order != "Primates" & order != "Perissodactyla")
par(mfrow=c(2,2))
plot(lm(order_size ~ gestation_mo, data=no_elephant_whale), which=2, xlab = "Gestation time")
plot(lm(order_size ~ afr_mo, data=no_elephant_whale), which=2, xlab = "Age of first reproduction")
plot(lm(order_size ~ litters_per_year, data=no_elephant_whale), which=2, xlab = "Litters per year")
plot(lm(order_size ~ life_mo, data=no_elephant_whale), which=2, xlab = "Life span")
dev.off()
```
